---
title: "Maddie Frick Assignment 4"
author: "MaddieFrick"
date: "11/19/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(effsize)
library(broom)
library(palmerpenguins)
library(ggbeeswarm)
library(kableExtra)
library(here)
library(lubridate)

rm(list=ls()) #clear the environment
```
## a) Introduction
This analysis includes the dataset from the Bonanza Creek LTER, part of the US Long Term Ecological Research Network about Snowshoe Hares, Lepus americanus. Snowshoe hairs are a keystone prey species in the northern boreal forests and their population flucutates over 8-11 year spans (Kielland, 1999). This dataset includes capture-recapture studies of hares at 5 locations in the Tanana valley between the years of 1999 to 2012. This exploratory report explores the relationship between the counts of hares per year, the impact of sex on the hare weight, and the relationship between hind foot length and hare weight.


## b) Data and Analyses
In this analysis, the amount of trappings of hares throughout 1999 to 2012 were inconsistent per year, with the maximum amount of trappings in 1999 with a total of 126, and a minimum amount in 2010 with two trappings. The inconsistent counts of hares per year could have an effect on the data analyzed. The sex of the hare significantly had an effect on the hare weight (p = 0.007093), but with only a small effect size (Cohen's d = -0.2904674). The hind foot length had a linear relationship with the hare weight with an R^2 value of 0.2988.


## c) Exploratory findings


### i) Annual juvenile hare trap counts

```{r}
hares<-read.csv(here("data","bonanza_hares.csv")) #read in the dataset
```


```{r}
#subset the data into just juveniles 
juvhare <- hares %>% #take the entire dataset
  filter(age=="j") %>% #select only the juvenile hares
  mutate(date = mdy(date)) %>% #the date is not in date format yet; this puts it in date format
  mutate(year = year(date)) #make a new column with only the year so that we can call the annual juv hare counts
```


```{r}
#count the total number of juveniles during each year
juvharebyyear <- juvhare %>% #take the juv dataset
  group_by(year) %>% #group them by the year column
  summarize(trappings = n()) #count the number of trappings per year

juvharebyyear$mean <- mean(juvharebyyear$trappings) #add the mean to the table
juvharebyyear$median <- median(juvharebyyear$trappings) #add the median to the table
juvharebyyear$max <- max(juvharebyyear$trappings) #add the max
juvharebyyear$min <- min(juvharebyyear$trappings) #add the min

```


```{r}
#create a finalized data visualization
ggplot(data = juvharebyyear, aes(x = year, y = trappings)) +
  geom_point(size = 2) +
  labs(x = "Year",
       y = "Trappings",
       title = "Annual Juvenile Hare Trappings",
       caption = "This graph shows a count of the total number of trapped hares each year between the years of 1999 to 2012.") +
  theme_light()
  

```

According to this graph, it looks like the amount of trappings per year decreased by 2001, and then fluctuated throughout the years of 2001-2012. The maximum number of trappings was in 1999 with a total of 126, and the minimum was in 2010 with two trappings. The mean number of trapped hares was 31.5 and the median was 18.5. There was a general trend of an initial decrease but not really any trend past 2001. I would suggest to do a more even number of hare trappings every year in order to standardize the juvenile hare population in future analyses. This would give a more accurate and consistent representation of the hare populations in every year, as a representative sample of 2 in 2010 will not cover the population very well. 


### ii) Visualize juvenile hare weights

Create finalized data visualization in which juvenile weights by sex and site are compared.


```{r}

labels <- c(f = "Female", m = "Male") #make a labels named vector that maps original names to new names for the facet
ggplot() +
  geom_beeswarm(data = juvhare, #make a beeswarm plot
                aes(x = grid, #x axis is the site, or grid
                    y = weight, #y axis is the weight 
                    ),
                size = 3,
                alpha = 0.6,
                pch = 16,
                color = "blue") +
  facet_wrap(~sex, labeller = labeller(sex = labels)) + #make three different graphs according to the sex
  labs(x = "Site",
       y = "Weight (g)",
       title = "Juvenile Hare Weights by Sex and Site") +
  theme_light() #add a theme!
```

This beeswarm plot tells me that the sex and site do not make much of a difference on the hare weights. In the bonrip site, the male hares seem to be a little bit heavier than the females in bonrip, but otherwise the weights are widespread at all three sites for both the male and the female hares. I am showing the audience the straight data because the summary statistics do not always tell all of the information. 


### iii) Juvenile weight comparison

```{r}
#take out the hares without the NA weights included
juvharewt <- juvhare %>% 
  filter(weight != "NA")

harewt <- juvharewt %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),#finds the mean of the size measurements
            median = median(weight, na.rm = TRUE), #same but ^
            sd = sd(weight, na.rm = TRUE), 
            n = n()) %>% 
  kable(col.names = c("Sex",
                      "Mean",
                      "Median", 
                      "Standard Deviation",
                      "Sample Size")) %>% 
  kable_styling()

```
This table shows the mean, median, standard deviation, and sample size of the juvenile male and female snowshoe hares included in this study.

```{r}
harewt
```


```{r, include = FALSE}
# Look at it:
# Histograms
ggplot(data = juvhare, aes(x = weight)) +
  geom_histogram(bins = 12) +
  facet_wrap(~sex)

# QQ Plots
ggplot(data= juvhare, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```


```{r, include = FALSE}
hare_f <- juvhare %>% #making a vector of the females
  filter(sex == "f") %>% 
  pull(weight)

hare_f<- hare_f[!is.na(hare_f)] #taking out the NA values

hare_m <- juvhare %>% #making a vector of the males
  filter(sex == "m") %>% 
  pull(weight)

hare_m<- hare_m[!is.na(hare_m)] #taking out the NA values


hare_ttest <- t.test(hare_f, hare_m) #run a t test
hare_ttest

hare_cohen <- cohen.d(hare_f, hare_m) #run the cohen's d test
hare_cohen
```
The actual difference in means of the male and female hares is 90.468. The mean body weight of the female hares was 855.39g and the males was 945.86g. I ran a two sample t test and cohen's d test on female and male hare weights. There was a significant different in body mass, and a small effect size (Cohen's d = -0.2904674). The resulting degrees of freedom was 325.02. The p value, the probability that we could have gotten sample means that are least as different as these if the null hypothesis (that they were pulled from populations with the same mean) was true is 0.007093. This means there is a .7% chance we could have gotten these from random chance if they were pulled from populations from the same mean, showing that sex probably has a significant effect on the weight of the hares. The cohen.d test showed us that the d estimate, the effect size, was -0.2904674, a small effect size of the sex of the hare on their weight.


### iv) Relationship between juvenile weight and hind foot length
```{r}
#explore the relationship
ggplot(data = juvhare, aes(x = hindft, y = weight)) +
  geom_point() + #the data looks somewhat linear
  labs(x = "Hind Foot Length (mm)",
       y = "Weight (g)",
       title = "Juvenile Weight and Hind Foot Length Relationship",
       caption = "This is an exploratory graph visualizing the relationship between juvenile hare weight and hind foot length" )
```
In this exploratory visualization, it looks like a linear relationship does make sense. The points are not exactly linear and the spread grows as the hind foot length increases, but the relationship is still somewhat linear, letting us use a linear model. Now I can model it using `lm()`.


```{r}
# create a linear model, stored as penguin_lm:
hare_lm <- lm(hindft ~ weight, data = juvhare)
# Return the complete overview:
summary(hare_lm)

hare_lm_tidy <- broom::tidy(hare_lm) #make a tidy table

# Get the intercept: 
hare_int <- hare_lm_tidy$estimate[1]
hare_int

# Then to get the flipper_length coefficient:
hare_coef <- hare_lm_tidy$estimate[2]
hare_coef

# Metrics at a glance: 
hare_lm_out <- broom::glance(hare_lm)
hare_lm_out

plot(hare_lm)

ggplot(data = juvhare, aes(x = hindft, y = weight)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() +
  ggpubr::stat_regline_equation(label.x = 180, label.y = 5700)

hare_cor <- cor.test(juvhare$hindft, juvhare$weight)
hare_cor

```

Simple linear regression was used to explore the relationship between hare foot length (mm) and weight (g) across both the male and female hares. A significant regression model was found ($\beta$ = `r round(hare_coef,3)`, F(`r hare_lm_out$df`,`r hare_lm_out$df.residual`) = `r round(hare_lm_out$statistic,1)`, p < 0.001) with an R^2^ of `r round(hare_lm_out$r.squared,3)`.The Multiple R^2^ value, the proportion of the variance in the dependent variable (weight) that was predictable from the independent variable (hind foot length) was 0.2988. This means that ~30% of the variance in the weight was predictable from the hind foot length. Both the intercept and the weight coefficients were significantly different than 0. The assumptions that we made doing this linear regression were linearity of the relationship between these factors, which was seen in our exploratory graph above, independence of errors, homoscedasticity of the errors, and normality of the error distribution. It would be better to include the hare sex as another variable because we know that the mean weights are different with male and female penguins, even though it was a small effect size. After using the plot() function, graphs 1 and 3 show us that the errors are not very homeodastic,the QQ plot showed us that the residuals were somewhat normal, and graph 4 showed Cook's distance. I would be a little bit concerned homoscedasticity. Here, we see that there is a strong positive correlation between hare hind foot length and weight (*r* = `r round(hare_cor$estimate,2)`, t(`r hare_cor$parameter`) = `r round(hare_cor$statistic,2)`, p < 0.001). 



Summary:
- ~30% of the variance in the weight was predictable from the hind foot length
- a significant regression model was found
- there was a positive correlation between hind foot length and weight

